---
title: "Global Disease Burden Analysis"
author: "REPLACE WITH THE AUTHOR NAMES" # AUTHOR NAMES HERE!
output: github_document
---

```{r include = F}
knitr::opts_chunk$set(echo = FALSE)


```

# Introduction

This document compares the Disability-Adjusted Life Year (DALY) diseases burden for three countries: COUNTRY 1, COUNTRY 2, and COUNTRY 3. We compare the burden due to communicable, maternal, neonatal, and nutritional diseases (CMNN), non-communicable diseases (NCDs), and overall disease burden for these countries. 

The source data is from the Institute for Health Metrics and Evaluation (IHME) Global Burden of Disease (GBD) study. 

```{r child='cmnn_burden.Rmd'}

if(!require(pacman)) install.packages(pacman)
pacman::p_load(tidyverse,
               here,
               gt)

cmnn_df <- 
  read.csv(here("data/burden-of-disease-cmnn.csv"))


names(cmnn_df)

cmnn_df <- 
  cmnn_df %>% 
  rename(DALY = DALYs..Disability.Adjusted.Life.Years....Communicable..maternal..neonatal..and.nutritional.diseases...Sex..Both...Age..Age.standardized..Rate.,
         Country = Entity)

unique(cmnn_df$Country)

#"Philippines" "Palestine" "South Africa" 

unique(cmnn_df$Year)

cmnn_PSE_ZAF_PHL <- 
  cmnn_df  %>% 
  filter(Country %in% c("Philippines","Palestine","South Africa" ))

cmnn_tab <- 
  cmnn_PSE_ZAF_PHL %>% 
  select(Country, Year, DALY) %>% 
  pivot_wider(names_from = Country, 
              values_from = DALY)


gt(cmnn_tab)

#plotting a line graph showing the cmnn trade 

cmnn_PSE_ZAF_PHL %>% 
  ggplot(aes(x = Year, y = DALY, fill = Country)) + 
  geom_col()
           

```

```{r child='ncd_burden.Rmd'}
```

```{r child='overall_burden.Rmd'}
```
