# Non-Communicable Disease Burden (NCD)

```{r}
# Load necessary packages
pacman::p_load(tidyverse, knitr, here)

# Import the non-communicable diseases data
data_ncd <- read_csv(here("data", "burden-of-disease-ncd.csv"))
```

## Table of Estimates for NCD Burden Over Time

```{r}
# Here render a table for the DALY burden over time for the three countries 
# You should pivot the data to show each country in a separate column.
# Use kable() from the knitr package to render the table.

library(dplyr)

filtered_data_ncd <- data_ncd %>%
  filter(Entity %in% c("Philippines", "Palestine", "South Africa")) %>%
  select(-Code) %>% 
  pivot_wider(names_from = Entity, values_from = DALY)

kable(filtered_data_ncd)

```

## Plot Showing Trends in NCD Burden Over Time

```{r}
# Here, plot a line graph showing the trend of DALY burden for the chosen countries over time. You can use ggplot2 to create this plot. Each country's line should be a different color.

filtered_data_ncdplot <- data_ncd %>%
  filter(Entity %in% c("Philippines", "Palestine", "South Africa")) %>%
  select(-Code)

ggplot(filtered_data_ncdplot, aes(x = Year, y = DALY, color = Entity)) +
  geom_line() +
  labs(title = "Trend of DALY for Philippines, Palestine, and South Africa")

```

## Summary of NCD Burden Findings

Provide a brief analysis based on the data presented in the table and chart. Highlight any significant findings or patterns. About 3 sentences.

Palestine and South Africa has a significant drop in DALYs through time, but for the Philippines it has an increase over the years until 2015.
